---
title: "FCC-BDC Availability Download"
author: "Christine Parker"
format: html
editor: visual
date: today()
---

# Set up environment

A [good reference](https://cran.r-project.org/web/packages/httr/vignettes/secrets.html) for different ways to keep things like usernames, passwords, and API tokens secure in your scripts.

```{r Load required libraries and settings}
library(httr)
library(jsonlite)
library(tidyverse)

# API base URL
base_url <- "https://bdc.fcc.gov/api/public/map/downloads"

# User credentials (replace with your username and token)
token <- Sys.getenv("FCC_API_TOKEN")
username <- Sys.getenv("FCC_USERNAME")

print(token)      # Should display your API token
print(username)   # Should display your username
```

# Function to get "as of" dates for availability data

This will create a function that after running will provide you with a list of the available data sets and their corresponding publication dates. Currently, there are "availability" and "challenge" data for download.

```{r create function to get as of dates}
get_as_of_dates <- function(username, token) {
  #create the url where you'll send your request
  url <- paste0("https://bdc.fcc.gov", "/api/public/map/listAsOfDates")
  
# Make the GET request, where your username and token are included in the header info to give you access
  res <- GET(
    url,
    add_headers(
      username = username,
      hash_value = token
    )
  )
  
##Could comment-out the section below if you don't want to have the debugging steps run   
# Check for successful response
  if (status_code(res) == 200) {
# Print the content of the response for debugging
    content_res <- content(res, "parsed")
    print(content_res)
    return(content_res)  # Return the parsed content
  } else {
# If not successful, print the error status code and message
    print(paste("Error:", status_code(res)))
    print(content(res, "text"))
    return(NULL)
  }
}

```

# Function to list availability data files for a specific date

This creates a function to download a list of all the availability files that are available for public download.

```{r function to list data files}
list_availability_files <- function(as_of_date) {
  url <- paste0(base_url, "/listAvailabilityData/", as_of_date)
  res <- GET(url, add_headers(username = username, hash_value = token))
  stop_for_status(res)
  content(res, as = "parsed", type = "application/json")


}
```

# Function to download a specific file

This function does the heavy lifting.

```{r debug version download}
download_file <- function(file_id, file_name, file_type = "1",
                          username = Sys.getenv("FCC_USERNAME"),
                          hash_value = Sys.getenv("FCC_HASH"),
                          download_path = "C:/Users/chris/ILSR Dropbox/Christine Parker/Christine Parker/RawData/BDC/BDC_J25") {

  if (!dir.exists(download_path)) {
    dir.create(download_path, recursive = TRUE)
  }

  sanitized_file_name <- gsub("[^A-Za-z0-9._-]", "_", file_name)
  full_file_name <- file.path(download_path, paste0(sanitized_file_name, ".zip"))

  url <- paste0("https://bdc.fcc.gov/api/public/map/downloads/downloadFile/availability/", file_id, "/", file_type)

  print(paste("File path:", full_file_name))
  print(paste("URL:", url))

  res <- httr::GET(
    url,
    httr::add_headers(
      username = username,
      hash_value = hash_value,
      `user-agent` = "Mozilla/5.0"
    ),
    httr::write_disk(full_file_name, overwrite = TRUE)
  )

  if (httr::http_error(res)) {
    stop(paste("Error downloading file_id:", file_id, "-", httr::http_status(res)$message))
  }

  message(paste("Downloaded:", full_file_name))
}


```

# Download Steps

This is the part of the script where you'll put all those functions to use! If you need to make edits to the categories of data you want to download, refer to the [API guide](https://us-fcc.app.box.com/v/bdc-public-data-api-spec) for more detail

```{r Download the data, include=FALSE, echo=FALSE}
# Step 1: Get "as of" dates
dates <- get_as_of_dates(username, token)

# Step 2: Select the latest date (or prompt user) - may need to edit number as the FCC is sequentially ordering the releases as of Nov 2024
latest_date <- dates$data[[7]]$as_of_date

# Step 3: List files for the selected date
availability_files <- list_availability_files(latest_date)
avail_data_list <- availability_files[["data"]] %>% map_dfr(~ .x) %>%
  #here I'm specifically interested in downloading the location-level data by state for only fixed broadband - you can choose other options - reference the API guide
  subset(category == "State" & subcategory == "Location Coverage" & technology_type == "Fixed Broadband" & file_type == "csv")


# Step 4: Download all files
# Iterate over all file_id values in avail_data_list
for (i in 1:nrow(avail_data_list)) {
  tryCatch({
    file_id <- avail_data_list$file_id[i]
    file_name <- avail_data_list$file_name[i]

    download_file(file_id, file_name, 
                  username = username, 
                  hash_value = token)

    Sys.sleep(0.5)  # prevent rate-limiting
  }, error = function(e) {
    message(paste("Error downloading file_id:", file_id, "-", e$message))
  })
}

message("All downloads attempted.")


```
